---
title: "EDA"
output:
  word_document:
    toc: yes
    fig_caption: yes
    keep_md: yes
  html_document: default
  pdf_document:
    toc: yes
editor_options: 
  chunk_output_type: inline
---

## Load the Required Packages

```{r eval=FALSE}
install.packages("pastecs")  
install.packages("fBasics")
install.packages("timeDate")
install.packages("timeSeries")
install.packages("stats")
install.packages("vars")
install.packages("ggfortify")
tinytex::install_tinytex()
install.packages("tibble")
```
## Load library

```{r}
library(pastecs)
library(timeDate)
library(timeSeries)
library(fBasics)
library(dplyr)
library(ggplot2)
library(tidyr)
library(vars)
library(ggfortify)
library(GGally)
library(ggpubr)
library(tinytex)
library(tibble)
```
## Read the data

```{r}
bud <- read.csv("~/PhD_Thesis/Data/budget_information.csv", header = TRUE)
bud <- as_tibble(bud)
```

## Inspect the data

```{r}
names(bud) # To know the variable names
str(bud)  # To know the class of the variables
```

## Computation of the annual budget performance

Annual Budget performance is Given as

$$bud\_perf=\frac{TEa}{TEe}*100\%$$

Where:
`bud_perf = Annual Budget Performance`
`TEa = Actual Total Expenditure`
`TEe = Estimated Total Expenditure`

```{r}
bud <- bud %>% 
  mutate(bud_perf=(TEa/TEe)*100)

names(bud)
```


## Transform the dataset into timeseries

```{r}
bud <- bud %>% 
  dplyr::select(-period) %>% 
  ts(start = 1999, end = 2020, frequency = 1)
```


## Summary Statistics
```{r}
pastecs::stat.desc(bud)
```


```{r}
fBasics::basicStats(bud)
```

## PLots
# Revenue variables with budget performance

We overlayed budget performace with other revenue variables to identify any visual relationship
Where:-
`IGRa` - Actual Internally Generated Revenue
`SAa`  - Actual Statutory Allocation
`VATa` - Actual VAT
`ORa`  - Actual Other Revenue
`CRA`  - Actual Capital Receipts
```{r}
bud_rev <- bud[,c(12,14,16,18,20,29)]
bud_rev %>% 
ggplot2::autoplot()
```


# Expenditure variables with budget performance

We overlayed budget performace with other expenditure variables to identify any visual relationship
Where:-
`PCa`    - Actual Personnel Cost
`OCa`    - Actual Overhead Cost
`CRFCa`  - Actual Consolidated Revenue Fund Charges
`SFa`    - Actual Stablisation Fund
`CAPEXa` - Actual Capital Expenditure
```{r}
bud_exp <- bud[,c(2,4,6,8,10,29)]
bud_exp %>% 
ggplot2::autoplot()
```


## Individual Graphs
# Revenue
```{r}
g_IGRa <- bud_rev[,1] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("Internally Generated Revenue")

g_SAa <- bud_rev[,2] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("Statutory Allocation")

g_VATa <-  bud_rev[,3] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("VAT")


g_ORa <- bud_rev[,4] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("Other Revenue")

g_CRa <- bud_rev[,5] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("Actual Capital Receipts")

g_bud_perf <- bud_rev[,6] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("Budget Performance")

ggarrange(g_IGRa, g_SAa, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)

ggarrange(g_VATa, g_ORa, 
          labels = c("C", "D"),
          ncol = 2, nrow = 1)

ggarrange(g_CRa, g_bud_perf, 
          labels = c("E", "F"),
          ncol = 2, nrow = 1)

```

# Expenditure

```{r}
g_PCa <- bud_exp[,1] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("Actual Capital Expenditure")

g_OCa <- bud_exp[,2] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("Actual Overhead Cost")

g_CRFCa <-  bud_exp[,3] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("Actual Consolidated Renue Fund Charges")


g_SFa <- bud_exp[,4] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("Actual Stablisation Fund")

g_CAPEXa <- bud_exp[,5] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("Actual Capital Expenditure")


ggarrange(g_PCa, g_OCa, 
          labels = c("G", "H"),
          ncol = 2, nrow = 1)

ggarrange(g_CRFCa, g_SFa, 
          labels = c("I", "J"),
          ncol = 2, nrow = 1)

ggarrange(g_CAPEXa, 
          labels = c("K"),
          ncol = 2, nrow = 1)

```


## Correlation Matrix

```{r}
bud_corr <-  bud[,c(2,4,6,8,10,12,14,16,18,20,29)]
bud_corr %>% 
  as.data.frame() %>% 
  ggpairs()
```