---
title: "Untitled"
output: html_document
---
---
title: "EDA"
output:
  word_document:
    toc: yes
    fig_caption: yes
    keep_md: yes
  html_document: default
  pdf_document:
    toc: yes
editor_options: 
  chunk_output_type: inline
---

## Load the Required Packages

```{r eval=FALSE}
install.packages("dplyr")

```
## Load library

```{r}
library(timeDate)
library(timeSeries)
library(fBasics)
library(dplyr)
library(ggplot2)
library(tidyr)
library(vars)
library(GGally)
library(ggpubr)
library(tinytex)
library(tibble)
library(tsibble)
library(tsibbledata)
library(stats)
library(feasts)
library(fabletools)
library(fpp3)
library(feasts)
library(fable)
library(fabletools)

```
## Read the data

```{r}
bud <- read.csv("~/PhD_Thesis/Data/budget_information.csv", header = TRUE)


```

## Inspect the data

```{r}
names(bud) # To know the variable names
str(bud)  # To know the class of the variables
```

## Computation of the annual budget performance

Annual Budget performance is Given as

$$bud\_perf=\frac{TEa}{TEe}*100\%$$

Where:
`bud_perf = Annual Budget Performance`
`TEa = Actual Total Expenditure`
`TEe = Estimated Total Expenditure`

```{r}
bud <- bud %>% 
  mutate(bud_perf=(TEa/TEe)*100)

names(bud)
```


## Transform the dataset into timeseries

```{r}
bud_seri <- bud %>% 
dplyr::select(-period) %>%
ts(start = 1999, end = 2020, frequency = 1)


bud_seri[,c(2:29)] <- round((bud_seri[,c(2:29)])/1e9,4) # Round values to Billion Naira


class(bud_series)
```


## Summary Statistics

```{r}
fBasics::basicStats(bud_series)
```

## PLots
# Revenue variables with budget performance

We overlayed budget performace with other revenue variables to identify any visual relationship
Where:-
`IGRa` - Actual Internally Generated Revenue
`SAa`  - Actual Statutory Allocation
`VATa` - Actual VAT
`ORa`  - Actual Other Revenue
`CRA`  - Actual Capital Receipts
```{r}
bud_rev <- bud_series[,c(12,14,16,18,20,29)]
bud_rev %>% 
ggplot2::autoplot()
```


# Expenditure variables with budget performance

We overlayed budget performace with other expenditure variables to identify any visual relationship
Where:-
`PCa`    - Actual Personnel Cost
`OCa`    - Actual Overhead Cost
`CRFCa`  - Actual Consolidated Revenue Fund Charges
`SFa`    - Actual Stablisation Fund
`CAPEXa` - Actual Capital Expenditure
```{r}
bud_exp <- bud_series[,c(2,4,6,8,10,29)]
bud_exp %>% 
ggplot2::autoplot()
```


## Individual Graphs
# Revenue
```{r}
g_IGRa <- bud_rev[,1] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("Internally Generated Revenue (N Billion)")

g_SAa <- bud_rev[,2] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("Statutory Allocation (N Billion)")

g_VATa <-  bud_rev[,3] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("VAT (N Billion)")


g_ORa <- bud_rev[,4] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("Other Revenue (N Billion)")

g_CRa <- bud_rev[,5] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("Actual Capital Receipts (N Billion)")

g_bud_perf <- bud_rev[,6] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("Budget Performance (%, Percentage)")

ggarrange(g_IGRa, g_SAa, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)

ggarrange(g_VATa, g_ORa, 
          labels = c("C", "D"),
          ncol = 2, nrow = 1)

ggarrange(g_CRa, g_bud_perf, 
          labels = c("E", "F"),
          ncol = 2, nrow = 1)

```

# Expenditure

```{r}
g_PCa <- bud_exp[,1] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("Actual Capital Expenditure (N Billion)")

g_OCa <- bud_exp[,2] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("Actual Overhead Cost (N Billion)")

g_CRFCa <-  bud_exp[,3] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("Actual Consolidated Renue Fund Charges (N Billion)")

g_CAPEXa <- bud_exp[,5] %>% 
       ggplot2::autoplot() +
         theme(text = element_text(size = 12, face = "bold"))+
         xlab("Period") + 
         ylab("Actual Capital Expenditure (N Billion)")


ggarrange(g_PCa, g_OCa, 
          labels = c("G", "H"),
          ncol = 2, nrow = 1)

ggarrange(g_CRFCa, g_CAPEXa, 
          labels = c("I", "J"),
          ncol = 2, nrow = 1)

```


## Correlation Matrix

```{r}
bud_corr <-  bud_series[,c(2,4,6,10,12,14,16,18,20,29)]
bud_corr %>% 
  as.data.frame() %>% 
  ggpairs()
```


## Box plot

```{r}
boxplot(bud_corr)
```

## Timeseries Model
# Time plot for budget performance
```{r}
ggplot2::autoplot(bud_corr[,10])+
  labs(y= "Budget Perfomance (% Percentange",
       x= "Year")
```




